---
title: "Find errors in data"
output: 
  rmarkdown::html_vignette: 
    df_print: kable
vignette: >
  %\VignetteIndexEntry{Find errors in data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(Ncpus = 1)
```


## Intro

Errorlocate uses validation rules from package `validate` to locate faulty
values in observations (or in database slang: erronenous _fields_ in _records_).

It follows this simple recipe (Felligi-Holt):

- Check if a record is valid (using supplied validation rules)
- If not valid then adjust the minimum number of values to make it valid.

`errorlocate` does this by translating this problem into a mixed integer 
problem (see other vignettes) and solving this mathematical problem.


## Methods

`errorlocate` has two main functions to be used:

- `locate_errors` for detecting errors
- `replace_errors` for replacing faulty values with `NA`

```{r setup}
library(validate)
library(errorlocate)
```

Let's start with a simple example:

We have a rule dat age cannot be negative:
```{r}
rules <- validator(age > 0)
```

And we have the following data set
```{r}
"age, income
 -10,    0  
  15, 2000
  25, 3000
  NA, 1000
" -> csv
d <- read.csv(textConnection(csv), strip.white = TRUE)
```

```{r, echo = FALSE}
d
```


```{r}
le <- locate_errors(d, rules)
summary(le)
```

`summary(le)` gives an overview of the errors found in this data set.
The complete error listing can be found with:

```{r}
le$errors
```

Which says that record 1 has a faulty value for age.


Suppose we expand our rules

```{r}
rules <- validator( r1 = age > 0
                  , r2 = if (income > 0) age > 16
                  )
```

With `validate::confront` we can see that rule `r2` is violated (record 2).

```{r}
summary(confront(d, rules))
```


What errors will be found by `locate_errors`?

```{r}
set.seed(1)
le <- locate_errors(d, rules)
le$errors
```

It now detects that `age` in observation 2 is also faulty, since it 
violates the second rule. Note that we use `set.seed`.
This is needed because in this example, either `age` or `income` can 
be considered faulty. `set.seed` assures that the procedure is 
reproducible.

With `replace_errors` we can remove the errors (which still need to be imputed).

```{r}
d_fixed <- replace_errors(d, le)
summary(confront(d_fixed, rules))
```
In which `replace_errors` set all faulty values to `NA`. 

```{r}
d_fixed
```

### Weights

`locate_errors` allows for supplying weigths for the variables. 
It is common that the quality of the observed variables differs.
When we have more trust in `age` we can give it more weight so it choose
income when it has to decide between the two (record 2):

```{r}
set.seed(1) # good practice, although not needed in this example
weight <- c(age = 2, income = 1) 
le <- locate_errors(d, rules, weight)
le$errors
```

For weights there are three different options:

- not specifying: all variables will have weight 1
- named vector: all records will have same set of weights
- named matrix, same dimension as the data: specify weights per record.

### Performance / Parallelisation

`locate_errors` solves a mixed integer problem. When the number of interactions between validation rules is large, finding an optimal
solution can be become computationally intensive. Both `locate_errors`
as well as `replace_errors` have a parallization option: `Ncpus` making
use of multiple processors. The `$duration` (s) property of each solution 
indicates the time spent to find a solution for each record. This can 
be restricted using the argument `timeout` (s).

```{r}
# duration is in seconds. 
le$duration
```
