---
title: "Data error!, but where?"
subtitle: "Errorlocate: find and replace erroneous fields in data using validation rules" 
author: "Edwin de Jonge"
date: "Statistics Netherlands / UseR! 2017"
output:
  beamer_presentation:
    includes:
      in_header: header.tex
    keep_tex: yes
    theme: "Frankfurt"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
library(errorlocate)
library(magrittr)
```

## {.plain .fragile}
\begin{centering}
  \includegraphics[width=0.8\paperwidth]{img/bad-data}
  \par
\end{centering}

## Data cleaning...

A large part of your job is spent in data-cleaning:

- getting your data in the right shape (e.g. `tidyverse`)

- assessing missing data (e.g. `VIM`)

- checking validity (e.g. `validate`)

- locating and removing errors: **`errorlocate`**!

- impute values for missing or erroneous data (e.g. `simputation`)

## {.plain}

\begin{center}
  \includegraphics[height=1\paperheight]{img/keep-calm-and-validate}
\end{center}

\usebackgroundtemplate{Hi}


## Validation rules?

Package `validate` allows to:

- formulate explicit data rule that data must conform to:

```{r}
library(validate)
check_that( data.frame(age=160, driver_license=TRUE), 
  age >= 0, 
  age < 150,
  if (driver_license == TRUE) age >= 16
)
```

## Explicit validation rules:

- Give a clear overview what the data must conform to.
- Can be used to reason about.
- Can be used to fix/correct data!
- Find error, and when found correct it.

### Note:

- Manual fix is error prone, not reproducible and not feasible for large data sets.
- Large rule set have (very) complex behavior, e.g. entangled rules: adjusting one value may 
invalidate other rules.

## Error localization

> Error localization is a procedure that points out fields in a data set 
that can be altered or imputed in such a way that all validation rules
can be satisfied.

## Find the error:

```{r}
library(validate)
check_that( data.frame(age=160, driver_license=TRUE), 
  age >= 0, 
  age < 150,
  if (driver_license == TRUE) age >= 16
)
```

It is clear that `age` has an erroneous value, but for more complex rule sets
it is less clear.

## Multivariate example:

```{r}
check_that( data.frame( age     = 3
                      , married = TRUE
                      , attends = "kindergarten"
                      )
          , if (married == TRUE) age >= 16
          , if (attends == "kindergarten") age <= 6
          )
```
Ok, clear that this is a faulty record, but what is the error? 

## Feligi Holt formalism:

> Find the minimal (weighted) number of variables that cause the invalidation of the data rules.

Makes sense! (But there are exceptions...)

Implemented in `errorlocate` (second generation of `editrules`).

## `errorlocate::locate_errors`

```{r, eval=TRUE}
locate_errors( data.frame( age     = 3
                  , married = TRUE
                  , attends = "kindergarten"
                  )
     , validator( if (married == TRUE) age >= 16
                , if (attends == "kindergarten") age <= 6
                )
     )$errors
```

## `errorlocate::replace_errors`

```{r, eval=TRUE}
replace_errors( 
    data.frame( age     = 3
              , married = TRUE
              , attends = "kindergarten"
              )
  , validator( if (married == TRUE) age >= 16
             , if (attends == "kindergarten") age <= 6
             )
)
```

## Internal workings:

`errorlocate`:

- translates error localization problem into a **mixed integer problem**, which
is solved with `lpsolveAPI`.

- contains a small framework for implementing your own error localization algorithms.


## Pipe friendly

The `replace_errors` function is pipe friendly:

```{r}
rules <- validator(age < 150)

data_noerrors <- 
  data.frame(age=160, driver_license = TRUE) %>% 
  replace_errors(rules)

errors_removed(data_noerrors) # contains errors removed
```

## Thank you!


\Large{Interested?}

```r
install.packages("errorlocate")
```

Or visit:

**[http://github.com/data-cleaning/errorlocate](http://github.com/datacleaning/errorlocate)**
