---
title: "Data errors, how to find them?"
subtitle: "Errorlocate: find and replace erroneous fields in data using validation rules" 
author: "Edwin de Jonge"
date: "Statistics Netherlands / uRos 2018"
output:
  beamer_presentation:
    includes:
      in_header: tex/header.tex
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
library(errorlocate)
library(magrittr)
```

## Who am I?

- Data scientist / Methodologist at Statistics Netherlands (aka CBS).
- Author of several R-packages, including `whisker`, `validate`, `errorlocate`, `docopt`, `daff`, `tableplot`,  `ffbase`, `chunked`, ...
- Co-author of _Statistical Data Cleaning with applications in R (2018)_ (together with @markvdloo)




## {.plain .fragile}
\begin{centering}
  \includegraphics[width=0.8\paperwidth]{img/bad-data}
  \par
\end{centering}

## Data cleaning...

A large part of your job is spent in data-cleaning:

- getting your data in the right shape (e.g. `tidyverse`, `dplyr`)

- assessing missing data (e.g. `VIM`, `datamaid`)

- checking validity (e.g. `validate`)

- locating and removing errors: **`errorlocate`**!

- impute values for missing or erroneous data (e.g. `simputation`, `recipes`)

## {.plain}

\begin{center}
  \includegraphics[height=1\paperheight]{img/keep-calm-and-validate}
\end{center}

## Statistical Value Chain

\begin{center}
 \includegraphics[width=\textwidth]{img/valuechain.pdf}
\end{center}

## Validation rules?

Package `validate` allows to:

- formulate explicit data rule that data must conform to:

```{r}
library(validate)
check_that( data.frame(age=160, driver_license=TRUE), 
  age >= 0, 
  age < 150,
  if (driver_license == TRUE) age >= 16
)
```

## Explicit validation rules:

- Give a clear overview what the data must conform to.
- Can be used to reason about.
- Can be used to fix/correct data!
- Find error, and when found correct it.

### Note:

- Manual fix is error prone, not reproducible and not feasible for large data sets.
- Large rule set have (very) complex behavior, e.g. entangled rules: adjusting one value may 
invalidate other rules.

## Error localization

> Error localization is a procedure that points out fields in a data set 
that can be altered or imputed in such a way that all validation rules
can be satisfied.

## Find the error:

```{r}
library(validate)
check_that( data.frame(age=160, driver_license=TRUE), 
  age >= 0, 
  age < 150,
  if (driver_license == TRUE) age >= 16
)
```

It is clear that `age` has an erroneous value, but for more complex rule sets
it is less clear.

## Multivariate example:

```{r}
check_that( data.frame( age     = 3
                      , married = TRUE
                      , attends = "kindergarten"
                      )
          , if (married == TRUE) age >= 16
          , if (attends == "kindergarten") age <= 6
          )
```
Ok, clear that this is a faulty record, but what is the error? 

## Feligi Holt formalism:

> Find the minimal (weighted) number of variables that cause the invalidation of the data rules.

Makes sense! (But there are exceptions...)

Implemented in `errorlocate` (second generation of `editrules`).

## `errorlocate::locate_errors`

```{r, eval=TRUE}
locate_errors( data.frame( age     = 3
                  , married = TRUE
                  , attends = "kindergarten"
                  )
     , validator( if (married == TRUE) age >= 16
                , if (attends == "kindergarten") age <= 6
                )
     )$errors
```

## `errorlocate::replace_errors`

```{r, eval=TRUE}
replace_errors( 
    data.frame( age     = 3
              , married = TRUE
              , attends = "kindergarten"
              )
  , validator( if (married == TRUE) age >= 16
             , if (attends == "kindergarten") age <= 6
             )
)
```

### Rule $r_i(x)$

A rule a disjunction of atomic clauses:

$$
r_i(x) = \bigvee_j C_i^j(x)
$$
with: 

$$
C_i^j(\la{x}) = \left\{
  \begin{array}{l}
 \la{a}^T\la{x} \leq b \\
  \la{a}^T\la{x} = b \\
  x_j \in F_{ij}  \textrm{with } F_{ij} \subseteq D_j \\
  x_j \not\in F_{ij} \textrm{with } F_{ij} \subseteq D_j \\
  \end{array}
\right.
$$


## Internal workings:

`errorlocate`:

- translates error localization problem into a **mixed integer problem**, which
is solved with `lpsolveAPI`.

- contains a small framework for implementing your own error localization algorithms.


## Pipe friendly

The `replace_errors` function is pipe friendly:

```{r}
rules <- validator(age < 150)

data_noerrors <- 
  data.frame(age=160, driver_license = TRUE) %>% 
  replace_errors(rules)

errors_removed(data_noerrors) # contains errors removed
```


## Interested?

\begincols
  \begincol{0.48\textwidth}
  \includegraphics[width=0.9\textwidth]{img/SDCR.jpg} 
  \endcol
  
  \begincol{0.48\textwidth}
  \begin{block}{SDCR}
M. van der Loo and E. de Jonge (2018)
\emph{Statistical Data Cleaning with applications in R}
Wiley, Inc.
\end{block}
\begin{block}{errorlocate}
\begin{itemize}
\item Available on \href{https://CRAN.R-project.org/package=errorlocate}{\underline{CRAN}}
\end{itemize}
\end{block}
\begin{block}{More theory?}
$\leftarrow$ See book 
\end{block}
\endcol
\endcols

Thank you for your attention (and enjoy The Hague)!
